---
title: "Lady Gaga's Discography"
author: "Daniël"
date: "2/13/2021"
output:
    flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE)
```

```{r}
library(tidyverse)
library(spotifyr)
library(plotly)
library(dplyr)

The_Fame <- get_playlist_audio_features("", "0tP9ws8v5IahnYpFWWZWjG")
Born_This_Way <- get_playlist_audio_features("", "7z51IAcHQtFa6DIsIQOGZE")
Artpop <- get_playlist_audio_features("", "0Cu6DccnrVil7ZDnE1bhwP")
Cheek_To_Cheek <- get_playlist_audio_features("", "7JEg2pZCBZktul0xqjDLiQ")
Joanne <- get_playlist_audio_features("", "0F9TvACmRg9jnWdTpHlPkM")
A_Star_Is_Born_Soundtrack <- get_playlist_audio_features("", "4jbR1SeCh5XpitDHxLTlB3")
Chromatica <- get_playlist_audio_features("", "0A3GuU5dqjCO4viWOZbfBZ")


```


# Lady Gaga: from The Fame to Chromatica
Lady Gaga's first album, The Fame, was released in August 2008 and immediately shot her to stardom. From that moment Gaga was on of the biggest artists in the world. Lady Gaga is one of the biggest pop artists in the world and is known for her musical range. After her debut album, she has released a variety of musical projects, ranging from mainstream pop to folk and jazz. I am very much

This aim of this website is to display my research on Lady Gaga's full discography. The albums I will compare are The Fame, Born This Way, Artpop, Cheek To Cheek (with Tony Bennett), Joanne, A Star Is Born Soundtrack and Chromatica. To study the musical differences, I will first rank the albums by commercial success by looking at the Spotify streams each album has. I will then look at the musical differences of each album and conclude if there is a relationship between the the commercial success and musical elements, such as energy and danceability.


# How do the albums score on the energy level?

Column {data-width=500}
-----------------------------------------------------------------------

### Chart A

```{r}
Lady_Gaga_Discography <- dplyr::bind_rows(
    The_Fame %>% mutate(category = "The_Fame"),
    Born_This_Way %>% mutate(category = "Born_This_Way"),
    Artpop %>% mutate(category = "Artpop"),
    Cheek_To_Cheek %>% mutate(category = "Cheek To Cheek"),
    Joanne %>% mutate(category = "Joanne"),
    A_Star_Is_Born_Soundtrack %>% mutate(category = "A_Star_Is_Born_Soundtrack"),
    Chromatica %>% mutate(category = "Chromatica")
  )

Lady_Gaga_Discography %>% ggplot(aes(x = valence, y = energy, color = "red")) + geom_jitter() + geom_smooth()

ggplotly()

```


Column {data-width=500}
-----------------------------------------------------------------------

### Chart B

```{r}
Lady_Gaga_Discography %>%
  ggplot(aes(x = energy, color = "red")) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~category)

ggplotly()
```


# Comparing the albums: does energy influence commercial success?
Chart B shows that there is a quite a big difference in energy between all the albums. Cheek To Cheek scores quite a bit lower than Chromatica for example, which is logical because the it is a jazz record consisting of more slower, chiller songs. Chromatica, on the contrary, has only upbeat songs and scores very high on the energy ranking. This rule seems to apply for all her pop albums. Artpop, The Fame, and Born This Way score much higher than Cheek To Cheek (jazz) and A Star Is Born. Joanne is kind of a mix, because it has both low and higher energy scores. The album consists of both upbeat and slower songs, as it is a mix of pop, folk/country and rock elements.

So altogether I am able to conclude that the more commercially successful albums are the ones who have more energetic songs. However, A Star Is Born is an outcast in this case, because it was very successful  even though it score higher in the middle of the energy ranking.
